---
title: "KNN regression"
author: "Ferran Ibáñez"
date: "15/02/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
The k nearest-neighbor estimator of $m(t) = E(Y |X = t)$ is defined as
$$\hat{m}(t)=\frac{1}{k} \sum_{i \in N_{k}(t)}y_i$$
where $N_k(t)$ is the neighborhood of $t$ defined by the $k$ closest points $x_i$ in the training sample.

We first read the Boston housing data from the library MASS.
```{r}
library(MASS)
data(Boston)
help(Boston)


x <- Boston$lstat
y <- Boston$medv
xy <- data.frame(x,y)
```

\section{Exercices}

\subsection{1. Write a function for computing the k-nn estimator of
 $m(t)$ for a given real value $t$}

The parameters used by the function are the df values from the training sample, the new observation $t$ and the $k$ number of nearest neighbours. We first compute the distance between all the points to the new one. Then we sort those distances and, finally, we compute the mean of the $y_i$ values based on the k nearest points to the new observation. The $C$ and $R$ parameters are used to grid the plots in case we want to compare the estimation using different k neighborhoods. 

```{r}
knn_regression<-function(t,df,k, C, R){
  # We create an empty df
  knn_reg <- data.frame(matrix(nrow = length(t), ncol = 2))
  
  # A list to save the different k's
  ks <- list()
  
  # K values:
  k_s <- k
  
  # A list of empty df
  for (i in k_s) {
    ks[[i]] <- knn_reg
  }
  # We generate the knn regressions:
  for (j in k_s) {
    for (p in t) {
      d_st_xy <- as.matrix(dist(c(p,df[,1])))[1,-1]
      d_st_xy_k <- sort(d_st_xy,partial=j)[j]
      N_st_k <- unname(which(d_st_xy <= d_st_xy_k))
      knn_reg[,1] <- seq(1,length(t))
      # Now we compute the knn estimator by the given definition of m(t):
      knn_reg[,2][p] <- sum(df[N_st_k,2])/j
    }
    # We save the estimated values in a list
    ks[[j]] <- knn_reg[,c(1, 2)]
  }
  # We generate different plots according to the k neighbourhoods:
  par(mfrow = c(R,C))
  for (z in k_s) {
    plot(df$y ~ df$x,
         data = Boston,
         cex = .8,
         col = "dodgerblue",
         main = paste0("k = ",z),
         ylab = "y (medv)",
         xlab = "x (lstat)")
    lines(ks[[z]][,1], ks[[z]][,2], col = "darkorange", lwd = 2.5)
    points(ks[[z]][,1], ks[[z]][,2], col = "yellow", lwd = 0.5)
  }
}
```



\subsection{3/4. Estimate $m(t[i])$ for $i=1,...,40$ using $k=50$. Plot $y$ against $x$. Then represent the estimated regression function.}

We just run the function we created before:

```{r}
knn_regression(1:40,xy,50, 1,1)
```


\subsection{5. Repeat the same exercise using different values of k.}


```{r,fig.width=10, fig.height=10}
k<-c(3,10,50,200)
knn_regression(1:70,xy,k,2,2)

```

